---
title: "ANCOVA - Mixed effects model"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
RIKZdata<-read.table("c:/liem/GEOG515/Zuur2009Datasets/RIKZ.txt",header = TRUE)
summary(RIKZdata)
```
##### Data: marine benthic data from nine inter-tidal areas along the Dutch coast collected by the Dutch institute RIKZ in the summer of 2002. In each inter-tidal area (denoted by ‘beach’), five samples were taken, and the macro-fauna and abiotic variables were measured. Zuur et al. (2007) used species richness (the number of different species) and NAP (the height of a sampling station compared to mean tidal level) 
```{r}
library(nlme)
# regression for each beach one by one
summary(lm(Richness ~ NAP, subset = (Beach==1), data=RIKZdata))
summary(lm(Richness ~ NAP, subset = (Beach==2), data=RIKZdata))
# regression for each beach in one for loop
Beta <- vector(length = 9)
for(i in 1:9){
  Mi <- summary(lm(Richness ~ NAP,subset = (Beach==i), data=RIKZdata))
  Beta[i] <- Mi$coefficients[2, 1]
}
Beta
```

```{r}
fExposure9 <- factor(c(0, 0, 1, 1, 0, 1, 1, 0, 0)) 
tmp2 <- lm(Beta ~ fExposure9)
summary(tmp2)
```
##### Mixed-effects model with random intercept
```{r}
RIKZdata$fBeach <- factor(RIKZdata$Beach)
Mlme1 <- lme(Richness ~ NAP, random = ~1 | fBeach, data = RIKZdata)
summary(Mlme1)
coef(Mlme1)
anova(Mlme1)
# 
Mlme0 <- lm(Richness ~ NAP + fBeach, data = RIKZdata)
summary(Mlme0)
anova(Mlme0)
```
##### Make graph
```{r echo=FALSE}
F0 <- fitted(Mlme1, level = 0)
F1 <- fitted(Mlme1, level = 1)
I <- order(RIKZdata$NAP)
NAPs <- sort(RIKZdata$NAP) 
plot(NAPs, F0[I], lwd = 4, type = "l",
ylim = c(0, 22), ylab = "Richness", xlab = "NAP")

for (i in 1:9){
x1 <- RIKZdata$NAP[RIKZdata$Beach == i] 
y1 <- F1[RIKZdata$Beach == i]
K <- order(x1)
lines(sort(x1), y1[K])
}
text(RIKZdata$NAP, RIKZdata$Richness, RIKZdata$Beach, cex = 0.9)
```
##### Mixed-effects model with random intercept and slope
```{r}
Mlme2 <- lme(Richness ~ NAP, random = ~1+NAP | fBeach, data = RIKZdata)
summary(Mlme2)
anova(Mlme2)
Mlme2 <- lme(Richness ~ 1, random = ~1+NAP | fBeach, data = RIKZdata)
summary(Mlme2)
anova(Mlme2)
```

```{r}
F0 <- fitted(Mlme2, level = 0)
F1 <- fitted(Mlme2, level = 1)
I <- order(RIKZdata$NAP)
NAPs <- sort(RIKZdata$NAP) 
plot(NAPs, F0[I], lwd = 4, type = "l",
ylim = c(0, 22), ylab = "Richness", xlab = "NAP")

for (i in 1:9){
x1 <- RIKZdata$NAP[RIKZdata$Beach == i] 
y1 <- F1[RIKZdata$Beach == i]
K <- order(x1)
lines(sort(x1), y1[K])
}
text(RIKZdata$NAP, RIKZdata$Richness, RIKZdata$Beach, cex = 0.9)
```
##### Random effects model

```{r}
Mlme3 <- lme(Richness ~ 1, random = ~1 | fBeach, data = RIKZdata)
summary(Mlme3)
anova(Mlme3)
```
