---
title: "PoiniPattern"
author: "LTT"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Spatial Point Patterns
The data for this R mark-up are on github folder. There is a file called north_derby_asthma.zip. Unzip it, and there will be four shapefiles inside.

```{r}
library(rgdal)
library(maptools)
data.dir <- "C:/Liem/GEOG515/Spring20/Datasets/north_derby_asthma"
# asthma point data
spasthma <- readOGR(dsn = data.dir, "spasthma")  
spasthma
summary(spasthma$Asthma)
# boundary layer
spbdry <- readOGR(dsn = data.dir, "spbdry")  
# source layer
spsrc <- readOGR(dsn = data.dir, "spsrc")  
# road layer
sproads <- readOGR(dsn = data.dir, "sproads")  

```

These data represent a random sample of children in the town of North Derby, Great Britain. There are children with asthma (cases) and children without (control). 
In addition we have the location of three large point sources of air pollutant and of motorways.
```{r}
# Plot the data in ggplot
library(ggplot2)
# Coerce all the data to data.frames for ggplot
ggbdry <- fortify(spbdry)
## Regions defined for each Polygons
ggasthma <- data.frame(long = coordinates(spasthma)[, 1], lat = coordinates(spasthma)[,  2], asthma = spasthma$Asthma)
ggsrc <- data.frame(long = coordinates(spsrc)[, 1], lat = coordinates(spsrc)[,  2], source = spsrc$Source)
ggroads <- fortify(sproads)

map <- ggplot(aes(x = long, y = lat), data = ggasthma)  #Initialize map with data
# Add points to map
map <- map + geom_point(aes(colour = asthma))
# Add boundary to map
map <- map + geom_path(aes(x = long, y = lat), data = ggbdry)
# Add roads to map, make them grey
map <- map + geom_path(aes(x = long, y = lat, group = group), data = ggroads, 
                       colour = "grey50")
# Add point sources to map, make them red
map <- map + geom_point(aes(x = long, y = lat), data = ggsrc, colour = "red", 
                        shape = 10, size = 4)
# And print the map to screen, making the axes equal along the way
map + coord_equal()
```

# Research question: Are the asthma cases clustered? 
Note: clustered relative to what? 
If you mean clustered geographically, so is the population (pay attention to the controls). 
What we really want to know is whether the cases are clustered relative to population (i.e. the controls). 
Furthermore, are the cases clustered around each other? or are they clustered around the point source?

# Analysis using spatstat

```{r}
library(spatstat)

# The main types of spatial data supported by spatstat are:
# + ppp point pattern 
# + owin window (spatial region) 
# + im pixel image 
# + psp line segment pattern 
# + tess tessellation 
# + pp3 three-dimensional point pattern 
# + ppx point pattern in any number of dimensions 
# + lpp point pattern on a linear network
# This emphasizes that the boundary definition is a fundamental part of point pattern analysis.
```


# Create the owin.  
```{r}
W <- as(spbdry, "owin") 
plot(W)
```

# Create the ppp

```{r}
# First way to create a ppp object: derive the cordinates then use them to create a ppp object
xy <- coordinates(spasthma)
pop.ppp <- ppp(x = xy[, 1], y = xy[, 2], window=W)
plot(pop.ppp)
## Warning: data contain duplicated points
# Ignore the duplicates warning
```

```{r}
# Second way to create a ppp object: convert a shapefile to a ppp object directly
asthma2 <- as(spasthma,"ppp")
asthma2
marks(asthma2)  <- NULL

Window(asthma2) <- W
plot(asthma2, main=NULL, cols=rgb(0,0,0,.2), pch=20)

```

# Quadrat density
You can compute the quadrat count and intensity using spatstat’s quadratcount() and intensity() functions. The following code chunk divides the state of Massachusetts into a grid of 3 rows and 6 columns then tallies the number of points falling in each quadrat.
The object Q stores the number of points inside each quadrat. You can plot the quadrats along with the counts as follows:

```{r}
Q <- quadratcount(asthma2, nx= 10, ny=6)
plot(asthma2, pch=20, cols="grey70", main=NULL)  # Plot points
plot(Q, add=TRUE)  # Add quadrat grid
```

# Compute the density for each quadrat

```{r}
Q.d <- intensity(Q)
# Plot the density
plot(intensity(Q, image=TRUE), main=NULL, las=1)  # Plot density raster
plot(asthma2, pch=20, cex=0.6, col=rgb(0,0,0,.5), add=TRUE)  # Add points
```

# Kernel density raster
The spatstat package has a function called density which computes an isotropic kernel intensity estimate of the point pattern. Its bandwidth defines the kernel’s window extent.
This next code chunk uses the default bandwidth.

```{r}
popden <- density(asthma2) # Using the default bandwidth
plot(popden, main=NULL, las=1)
contour(popden, add=TRUE)
```


# Kernel density adjusted for covariate
# ρ function
Under regularity conditions, ρ is proportional to the ratio of two probability densities, the numerator being the density of
covariate values at the points of the point process, while the denominator is the density of covariate values at random locations in space.
Asthma case intensity is estimated following the population density raster. The outputs include a plot of ρ vs. population density and a raster map of ρ controlled for population density.

```{r}
# Compute rho using the ratio method
asthma3 <- as(spasthma,"ppp")
asthma_dat <- pop.ppp[asthma3$marks$Asthma == "case", ]
asthma_dat

asthma_den <- density(asthma_dat) # Using the default bandwidth
plot(asthma_den, main=NULL, las=1)
contour(asthma_den, add=TRUE)

# rhohat: Relative intensity as a function of covariate
rho <- rhohat(asthma_dat, popden,  method="ratio")
# Generate rho vs covariate plot
plot(rho, las=1, main=NULL, legendargs=list(cex=0.8, xpd=TRUE, inset=c(0, 0) ))
```
The following code chunk generates the map of the predicted density when controlled for the population density. If the covariate does a good job in explaining the asthma density, the predicted density map should be nearly uniform across the map.

```{r}
# plot density map of asthma cases predicted by poppulation density
pred <- predict(rho)
cl   <- interp.colours(c("lightyellow", "orange" ,"red"), 100) # Create color scheme
plot(pred, col=cl, las=1, main=NULL)

# Now plot density map of asthma cases (Note: the density fuction has many parameters. Go to the manual of the spatstat package to learn the details)
asthma_dat <- pop.ppp[asthma3$marks$Asthma == "case", ]
aden <- density(asthma_dat, sigma=0.05)
plot(aden, main=NULL, las=1)
contour(aden, add=TRUE)

# Scatter plot of two density map
pairs(pred, aden)
```



## K and L functions
# Plot K function
```{r}
K_asthma <- Kest(pop.ppp,correction=c("border", "isotropic", "rigid", "translate"))
plot(K_asthma)
```

# envelope: simulation envelopes for a summary function or compute simulation envelopes for a ﬁtted model

```{r}
env <- envelope(Y = pop.ppp, fun = Kest, nsim = 29)
plot(env)
```

# Plot L function
```{r}
L_asthma <- Lest(pop.ppp,correction=c("border", "isotropic", "rigid", "translate"))
plot(L_asthma)
```

# envelope: simulation envelopes for a summary function or compute simulation envelopes for a ﬁtted model

```{r}
env <- envelope(Y = pop.ppp, fun = Lest, nsim = 29)
plot(env)
```

# K function with marked point pattern

```{r}
marks(pop.ppp) <- spasthma$Asthma
# How many point are there?
summary(pop.ppp)
```


# There are  215 asthma cases out of 1291 points.  What is the K-function of the asthma cases?
```{r}
K_dat <- Kest(pop.ppp[marks(pop.ppp) == "case", ],correction=c("border", "isotropic", "rigid", "translate"))
plot(K_dat)
L_dat <- Lest(pop.ppp[marks(pop.ppp) == "case", ],correction=c("border", "isotropic", "rigid", "translate"))
plot(L_dat)
# It seems the asthma cases were clustered. Is it truely clustered or due to the clustered pattern of the population?
# Need to take into account of the population density
```

# Create storage space for the simulations
```{r}
r <- seq(0, 0.2, by = 0.01)
K_dat <- Kest(pop.ppp[marks(pop.ppp) == "case", ], r = r)
K.storage <- matrix(NA, nrow = length(r), ncol = 40)
K.storage[, 1] <- K_dat$border
for (k in 2:40) {
    # randomly create 215 'cases'
    sim.case <- sample(1291, 215)
    K.storage[, k] <- Kest(pop.ppp[sim.case, ], r = r)$border
}

```

# use the adply function to calcualte the quantile of each row
# Extract the 5th and 95th quantiles of each row to create a 90% confidence interval

```{r}
library(plyr)  
K.env <- adply(K.storage, 1, quantile, c(0.05, 0.95), na.rm = TRUE)
plot(r, K_dat$border, type = "l")
lines(r, K.env[, 2], lty = 2)
lines(r, K.env[, 3], lty = 2)
# Comment: There is not really any significant evidence of clustering beyong that of the population. If you were to randomly pick 215 cases from the population, you could get a K-function just like the K function of the asthma cases.

```

# Regression with marked point patterns
Each point has a mark: case or control. Are points more likely to be cases if they are near roads? Near point sources? This is a question for logistic regression.

``` {r}
asthma.df <- spasthma@data
class(asthma.df)
asthma.df$x <- coordinates(spasthma)[, 1]
asthma.df$y <- coordinates(spasthma)[, 2]
asthma.df$smoking <- asthma.df$Nsmokers > 0
# There is some missing gender data... remove those
model <- glm(Asthma ~ d2source1 + d2source2 + d2source3 + smoking + HayFever + 
               Age + Gender + roaddist2, data = asthma.df, family = binomial, subset = Gender != 0)
summary(model)
# There appears to be a weak effect due to the point sources.  
# Question: Is there a residual spatial effect? How to handle such potential issue? Later.
```


# Average nearest neighbor analysis
Next, we’ll compute the average nearest neighbor (ANN) distances between asthma cases.

``` {r}
# ASthma cases only
asthma_dat <- pop.ppp[asthma3$marks$Asthma == "case", ]
# Identify nearest neighbours: nnwhich(X) 
nearest_neighbors<- nnwhich(asthma_dat) 
# To compute the average first nearest neighbor distance, set k=1
mean(nndist(asthma_dat, k=1))
# To compute the average second nearest neighbor distance, set k=2
mean(nndist(asthma_dat, k=2))

```

The average nearest neighbor function can be expended to generate an ANN vs neighbor order plot. In the following example, we’ll plot ANN as a function of neighbor order for the first m closest neighbors:
```{r}
m<-30
ANN <- apply(nndist(asthma_dat, k=1:m),2,FUN=mean)
plot(ANN ~ eval(1:m), type="b", main=NULL, las=1)
```

# Test of CSR for nearest-neighbor distance

```{r}
# clark-Evans test
clarkevans.test(asthma_dat, correction="none", alternative="clustered")
# Hopkins-Skellam test
hopskel.test(asthma_dat, correction="none", alternative="clustered")

```

# Nearest-neighbour function G and empty-space function F
Empty-space function F: Distance from a fixed location u ∈ R2 to the nearest point of the process: ‘empty-space distance’(or ‘spherical contact distance’ or ‘void distance’)

```{r}
# Estimates of F(r) and G(r) are computed from a point pattern X by the commands Fest(X) and Gest(X), respectively. The syntax for these commands is very similar to that of Kest.
Fs_asthma_dat <- Fest(asthma_dat)
Gs_asthma_dat <- Gest(asthma_dat)
plot(Fs_asthma_dat)
plot(Gs_asthma_dat)

```

