---
title: "GEOG 515 - spring 15: Introduction to R"
output: html_document
---

This is an introduction to R. It describes basic operations of data input/output, making graphs and simple statistics.

## Getting Start

Some of R packages (i.e., libraries) are part of the basic installation. Others can be downloaded from CRAN  which currently hosts over 1000 packages for various purposes. If you want to list all installed package, type:

#### library()

A package is loaded into R using the library command, so to load the "diagram" package you should
enter:

#### library(diagram)

To list of functions in the diagram package with brief descriptions, enter: 

#### help(package=diagram)

To view the vignette, type:

#### vignette("diagram")

To look at the documentation for the function plotmat(), type:

#### ??plotmat

### Examples of function option()

To view defaults, type:

#### options() 

To know the default of the number of digits, type:

#### getOption("digits")

If you want to change the maximum number of digits printed from 7 (default) to 10, type:

#### options(digits=10)

### Using R as a calculator

```{r}
sqrt(16)+6^3
a=12^2
a
```

### Vectors

```{r}
# Examples of creating vectors:
v1 <-  c(2.5, 4,  7.3, 0.1)
v2 <-  c("A", "B", "C", "D")
v3 <-  -3:3
v4<- seq(0,   2,  by=0.3)
v5<- seq(0,   2,  len=6)
v6<- rep(1:5, each=2)
v7<- rep(1:5,  times=2)
v1 
v2 
v3 
v4 
v5 
v6 
v7
```

#### Examples of referencing vector elements

```{r}
x <-  c(4, 9,  2,  12,  11,  3)
x[4]
x[1:3] 
x[c(2,5,6)]
x[-3]
x[-c(4,5)] 
x[x>4] 
x
```

#### Examples of vector operations:

```{r}
x <-  c(2, 4, 6, 0)
y <-  x^2 + 1
y
x*y
```

#### Example of matrix:

```{r}
x <-  matrix(c(1,2,3,4,5,6), nrow=2,  ncol=3, byrow=TRUE, dimnames=list(rows=c("row1", "row2"), cols=c("col1", "col2", "col3")))
x
```

#### Example of list:

```{r}
z <-  list(c(1,2,3), "Alisa", x)
z
z <-  list(num=c(1,2,3), name="Alisa", account=x)
z
```

#### Example of dataframe

```{r}
data()
data  <-  data.frame(chickwts) 
data
head(data)       # Print the  first few rows
tail(data)  # Print the  last few rowr
names(data) 	# Column names colnames(data); rownames(data)    # Column and row names 
dim(data)	# Dimension of  the  dataframe
data[ ,c("weight",  "feed")]	# "weight" and "feed"  columns 
data$weight 	# Get the  column "weight"
data[1:5, ]	# Get the  first five rows
```

#### # Example of importing data into R

```{r}
setwd("C:/Liem/GEOG515/Spring15/Labs/")
getwd()
data  <-  read.table("DuhramNCdata.txt", header=TRUE,  sep="\t")
str(data) # Gives  the  structure of  data

# Load data  using  file.choose()
# data  <-  read.table(file.choose(), header=TRUE,  sep="\t")

# Load a csv file:
# read.csv(file.choose(), dec='.', sep=',')
```

### Simple summary statistics

```{r}
dry <- c(77, 93, 92, 68, 88, 75, 100)
dry
sum(dry)
length(dry)
mean(dry)
sum(dry)/length(dry) ## Checking
median(dry)
sd(dry)
var(dry)
sd(dry)^2
sum((dry-mean(dry))^2) / (length(dry)-1) ## Checking
min(dry)
max(dry)
summary(dry)
```

#### Basic  Scatterplot 

```{r}
x <-  c(1:20)
y <-  c(1:10,9:0)
plot(x, y)
```

#### Add some features:

```{r}
plot(x,  y, xlab="X label", ylab="Y label", main="my first graph in R",  pch=15,  col="red")
```

#### And more features:

```{r}
plot(x, y,  main = "My Chart Title", type = "b", xlim = c(0,20), 
     ylim = c(0,11), xlab ="X", ylab = "Y", pch=16,   col = "red",       
     bty = "n", xaxs="i", yaxs = "i",	las=1) 
# Add text:
text(6,7, "My text", col = "green" ) 

# Distinguish between two separate  groups 
plot(x,  y, xlab="X", ylab="Y", main="Y vs  X", 
     pch=ifelse(x%%2==1, 5,  19), 
     col=ifelse(x%%2==1, "red", "blue"))

## Example of 0uter Margin annotation 
my_date <- format(Sys.time(), "%m/%d/%y") 
my_text <-"manuscript 1"
mtext(my_text, side = 1, line = .75, cex=0.7, outer = T, adj = 0) 
mtext(my_date, side = 1, line =.75, cex = 0.7, outer = T, adj = 1) 
```

#### Basic  Line  Graphs

```{r}
plot(x, y, type="l", lty=2, lwd=2,  col="blue")

# The  lines argument can be,  (1) two separate  vectors
# where one vector is the  x-coordinates and the  other   is the
# y-coordinates (2) a two-column  matrix or  (3) a two-element  list
# with  x and y components. 
plot(x,  y, type="n") 
lines(x, y, type="b")

plot(x,  y, type="n") 
lines(cbind(x,y),  type="l", lty=1, col="blue")

# If there   is only  one component then  the  argument is plotted against
# its index  (same with  plot and points)
plot(sort(x),  type="n")
lines(x, type="b", pch=8,  col="red")
lines(y,  type="l", lty=6, col="blue")
```

#### Example of adding lines in plot area

```{r}
x <- 0:100
y <- x^2
x_pts <- c(15, 35, 35, 15, 15)
y_pts <- c(8300, 8300, 6500, 6500,8300)
plot(x,y, type = "n", main="Add Lines, Points, Arrows Examples")
points(x, y,col = "black", lty=1, type = "l")
grid(col="lightgrey", lty=1)
abline(h=2500, col="red")
abline(v=50, col = "blue")
arrows(65,2400, 85,2400, code=3,  col="orange", length = 0.1)
lines(x_pts, y_pts, lty=1, col = "green")
points(x_pts,y_pts, type = "p",  col ="red",pch=19)
```

### More Graphical Tools with R

```{r}
# now we use the trees dataset in R
data<- trees
data
```

The trees dataset has three fields: Girth, Height, and Volume.
There are different ways to use the fields of the trees dataset:
.	Attach the dataset as a data frame:  attach(trees)	(we learned this way last week)
.	Use the "data=trees" syntax, 
.	Wrap the command in the "with" command, or
.	Use the syntax dataset$field for each variable.

The following commands will produce the same result:

```{r}
#Pay attention to the labels of the x and y axes.
plot(Height~Girth, data=trees)       # you do not need to "attach" "trees" before this command
with(trees, plot(Height~Girth))   # you do not need to "attach" "trees" before this command
plot(trees$Girth, trees$Height) 
```


Now try this command:
```{r}
plot(Height ~Girth, data=trees, xlab= "GIRTH", ylab= "HEIGHT",main=  "Scatter Plot\n Height vs. Girth of Black Cherry Trees")
# Compare the labels with those in the previous plot and pay attention to the plot title. (\n: go to the next line).
```

Now we use the swiss dataset. Try and observe the output:

```{r}
swiss
plot(swiss)
```

Now try and compare the two graphs below:
```{r}
plot(~ Fertility + Education + Catholic, data = swiss)
```

```{r}
pairs(~ Fertility + Education + Catholic, data = swiss)
```

Try these to see how to add a fitted curve to a scatter plot:
```{r}
plot(trees$Height,trees$Girth)
 lines(lowess(trees$Girth~trees$Height))
```
Plot two scatter plots on the same graph:
```{r}
plot(trees$Volume~trees$Height,col="green",xlab="Height",ylab="Girth/Volume")
points(trees$Girth~trees$Height,col="red")   # "points" adds a new series to an existing graph
```

### Barplot

Now we use the VADeaths dataset (Death Rates in Virginia (1940)). Type the following functions and compare the graphs (delete each graph before typing the next function.
```{r}
barplot(VADeaths, main="Death Rates in Virginia (1940)")
```

```{r}
barplot(VADeaths,col=rainbow(5), main="Death Rates in Virginia \n(1940)")
```

```{r}
barplot(VADeaths,col= heat.colors(5), main="Death Rates in Virginia \n(1940)")
```

```{r}
barplot(VADeaths,beside=TRUE,col=rainbow(4))
legend("topleft",  cex=1.0,rownames(VADeaths),bty="n", fill=rainbow(4))
```
Delete the graph and try:
```{r}
barplot(VADeaths, beside=TRUE)
legend("topleft",  cex=0.7,rownames(VADeaths),bty="y", fill=rainbow(4))
title(main = list("Death Rates in Virginia (1940)", font = 4)) 
```

### Boxplot

A box plot is the standard box and whiskers plot. Usage:  boxplot(x, ...) 
x can be a vector or a formula (y ~ grp), where grp is a factor. 
Now we use the chickwts dataset with two fields: weight and feed. 

```{r}
boxplot(chickwts$weight,col="blue")
boxplot(weight ~ feed, data=chickwts, col="red")

```

```{r}
boxplot(weight ~ feed, data=chickwts, col=rainbow(4), horizontal=TRUE)

```

### Histogram

The function hist() computes a histogram of the given data values: hist(x, ...)
x is the vector of values to be in the histogram. 

```{r}
hist(chickwts$weight, plot=FALSE)
hist(chickwts$weight, plot=TRUE)

h<- hist(chickwts$weight, plot=FALSE)

# compute the frequency polygon
diffBreaks <- h$mids[2] - h$mids[1]
xx <- c( h$mids[1]-diffBreaks, h$mids, tail(h$mids,1)+diffBreaks ) 
yy <- c(0, h$density, 0) 
# draw the histogram
hist(chickwts$weight, prob = TRUE, xlim=range(xx), border="gray", col="green",xlab="weight",main="Histogram of Chicken Weight") 
# adds the frequency polygon
lines(xx, yy, lwd=2, col = "blue")
# adds the densily curve if you want
lines(density(chickwts$weight),col="red")

```

###Pie charts

function pie(x, ...) 
Example:

```{r}
sales <- c(0.12, 0.3, 0.26, 0.16, 0.04, 0.12) 
names<- c("Blueberry", "Cherry", "Apple", "Boston Cream", "Other", "Vanilla Cream")
pie(sales,labels=names,col=rainbow(length(names)), clockwise=TRUE)
```

### Line Chart

We will use the Orange dataset in R

```{r}
Orange
# convert factor to numeric for convenience 
Orange$Tree <- as.numeric(Orange$Tree) 
ntrees <- max(Orange$Tree)

# get the range for the x and y axis 
xrange <- range(Orange$age) 
yrange <- range(Orange$circumference) 

# set up the plot 
plot(xrange, yrange, type="n", xlab="Age (days)",
    ylab="Circumference (mm)" ) 
colors <- rainbow(ntrees) 
linetype <- c(1:ntrees) 
plotchar <- seq(18,18+ntrees,1)

# add lines 
for (i in 1:ntrees) { 
  tree <- subset(Orange, Tree==i) 
  lines(tree$age, tree$circumference, type="b", lwd=1.5,
    lty=linetype[i], col=colors[i], pch=plotchar[i]) 
} 

# add a title and subtitle 
title("Tree Growth", "(created by Pokemon)")

# add a legend 
legend(xrange[1], yrange[2], 1:ntrees, cex=0.8, col=colors,
  	pch=plotchar, lty=linetype, title="Tree")
```


