---
title: "spatialRegression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bring in needed packages

```{r}
library(spdep)
library(rgdal) 
library(rgdal)
library(Hmisc)
library(car)
library(AER)
library(spatialreg)
library(spgwr)
# require(simrel)
# require(ggplot2)
# require(reshape2)
# library(HistogramTools)
```

# Bring in data. Change directory as needed
```{r}
# setwd("C:/Liem/GEOG515/Spring20/Labs/")
# UScounties <- readOGR(".", "tl_2019_us_county")
# summary(UScounties)
# spplot(UScounties, "ALAND")
```

```{r}
setwd("C:/Liem/GEOG515/Spring20/Labs/")
NY8 <- readOGR(".", "NY8_utm18")
summary(NY8)
spplot(NY8, "PCTOWNHOME")
spplot(NY8, "PCTAGE65P")
spplot(NY8, "PEXPOSURE")
spplot(NY8, "Z")
```

# Test normality
```{r}
shapiro.test(NY8$PEXPOSURE)
shapiro.test(NY8$PCTAGE65P)
shapiro.test(NY8$PCTOWNHOME)
shapiro.test(NY8$Z)
```

# Explore histograms
```{r}
par(mfrow=c(2,2))
hist(NY8$PEXPOSURE,breaks=20, col="red", xlab="PEXPOSURE", main="") 
hist(NY8$PCTAGE65P,breaks=20, col="red", xlab="PCTAGE65P", main="") 
hist(NY8$PCTOWNHOME,breaks=20, col="red", xlab="PCTOWNHOME", main="") 
hist(NY8$Z,breaks=20, col="red", xlab="Z", main="") 
```

# Construct neighbor matrix
```{r}
NY_nb.queen.false <- poly2nb(NY8, queen = FALSE)
summary(NY_nb.queen.false)
NY_nb.queen.true <- poly2nb(NY8, queen = TRUE)
summary(NY_nb.queen.true)
```

# Construct row-standardized spatial weights
```{r}
NY_W <- nb2listw(NY_nb.queen.false, style = "W", zero.policy = TRUE)
summary(NY_W)
# Look at one object's weights
NY_W$weights[[1]]
# Construct unstandardized spatial weights
NY_B <- nb2listw(NY_nb.queen.false, style = "B", zero.policy = TRUE)
# Look at one object's weights
NY_B$weights[[1]]
#
```

# Moran's I tests on Z
```{r}
moran.test(NY8$Z, NY_W)
moran.test(NY8$Z, NY_B)
#
```

# SAR on Z (deprecated in spdep but moved to the spatialreg package)
```{r}
NY.SAR1<-lagsarlm(Z ~ 1, data=NY8, listw=NY_W,type="lag")
summary(NY.SAR1)
moran.test(resid(NY.SAR1), NY_B)
```

# SAR on Z (deprecated in spdep but moved to the spatialreg package)
```{r}
NY.SAR2 <- spautolm(Z ~ 1, data=NY8,listw=listw_NY, family="SAR", method="eigen")
summary(NY.SAR2)
moran.test(resid(NY.SAR2), NY_B)
```

# CAR on Z (deprecated in spdep but moved to the spatialreg package)
```{r}
NY.CAR <- spautolm(Z ~ 1, data=NY8,listw=listw_NY, family="CAR", method="eigen")
summary(NY.CAR)
moran.test(resid(NY.CAR), NY_B)
AIC(NY.SAR2,NY.CAR)
BIC(NY.SAR2,NY.CAR)
```

# SMA on Z (deprecated in spdep but moved to the spatialreg package)
```{r}
NY.SMA <- spautolm(Z ~ 1, data=NY8,listw=listw_NY, family="SMA", method="eigen")
summary(NY.SMA)
moran.test(resid(NY.SMA), NY_B)
```

# spatial lag model (in spatialreg package)

```{r}
Model_2<-lagsarlm(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME , data=NY8, listw=NY_W,
                   method="eigen", quiet=FALSE)
rho.hat3 <- Model_2$rho
moran.test(resid(Model_2),NY_W)
Model_2.fit<-fitted(Model_2)
Model_2.resid<-resid(Model_2)
shapiro.test(Model_2.resid)
summary(Model_2)
Model_2_s<-summary(Model_2)
Model_2_s$coefficients[2]

```

# spautolm's spatial error model (in spatialreg package)
```{r}
Model_1<-errorsarlm(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME , data=NY8, listw=NY_W,
                   method="eigen", quiet=FALSE)
summary(Model_1)

lambda.hat2 <- Model_1$lambda
moran.test(resid(Model_1),NY_W)
Model_1.fit<-fitted(Model_1)
Model_1.resid<-resid(Model_1)
shapiro.test(Model_1.resid)
Model_1_s<-summary(Model_1)
Model_1_s$Coef[,]
```


# GWR (in spgwr package)
```{r}
NY.lm <- lm(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME , data=NY8)
summary(NY.lm)
NY.bw <- gwr.sel(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME , data=NY8,
  coords=cbind(NY8$X, NY8$Y))
NY.gauss <- gwr(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME , data=NY8,
  coords=cbind(NY8$X, NY8$Y), bandwidth=NY.bw, hatmatrix=TRUE)
NY.gauss
NY.d <- gwr.sel(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME , data=NY8,
  coords=cbind(NY8$X, NY8$Y), gweight=gwr.bisquare)
NY.bisq <- gwr(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME , data=NY8,
  coords=cbind(NY8$X, NY8$Y), bandwidth=NY.d, 
  gweight=gwr.bisquare, hatmatrix=TRUE)
NY.bisq$SDF$PEXPOSURE

spplot(NY.bisq$SDF, "PEXPOSURE")

```


